---
title: "Part2: using APIs and making data visualizations"
format: html
editor: visual
---

```{r}
if (!require("tidycensus", quietly = TRUE)) {
    install.packages("tidycensus", repos = "http://cran.us.r-project.org")
}
```

```{r}
library(tidycensus)
library(here)
```

## Some introduction

```{r}
library(tidycensus)
library(tidyverse)

## Save API Key in the global environment
#census_api_key("d42ebf0b6ff84c585897e8cf49ae6a0fc98febdd", install = TRUE)
## Activate API key
readRenviron("~/.Renviron")
```

#### Check the label list in acs

```{r}
# check
v_acs20 <- load_variables(2020, "acs5", cache = TRUE)
var_list <- c('B19013_001', 'B19201_001', 
              'B08303_001', 'B08134_101', 'B08134_111',
              'B08301_016', 'B08301_017', 'B08301_018',
              'B08301_019', 'B08301_020', 'B09002_001')

  # B19013_001: MEDIAN HOUSEHOLD INCOME
    # B19013_001E: Median Household Income
    # B19013_001M: Median Household Income Margin of Error
  # B19201_001: NONFAMILY HOUSEHOLD INCOME
  # B08303_001: travel time to work
  # B08134_101: Estimate!!Total:!!Walked: MEANS OF TRANSPORTATION TO WORK BY TRAVEL 
  # B08134_111: Estimate!!Total:!!Taxicab, motorcycle, bicycle, or other means:
  # B08301_016: Estimate!!Total:!!Taxicab
  # B08301_017: Estimate!!Total:!!Motorcycle
  # B08301_018: Estimate!!Total:!!Bicycle
  # B08301_019: Estimate!!Total:!!Walked
  # B08301_020: Estimate!!Total:!!Other means
  # B09002_001: Estimate!!Total: OWN CHILDREN UNDER 18 YEARS BY FAMILY
```

# Data Analysis

### Research Question

-   The median income distribution by different commuting time in Texas.
-   The association between commuting time with different means of transportation and median income or median non-family income.
-   Try to use Bayesian kernel machine regression to explore unlinear assocition between multivariables and median income.

### Original Data

-   Data was downloaded from TidyTuesday. The data this week comes from many different sources but originally came from the US Department of Education. Tuition and fees by college/university for 2018-2019, along with school type, degree length, state, in-state vs out-of-state from the Chronicle of Higher Education. [Data Source](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-03-10/readme.md)

### Data Dictionary

-   Here is a data dictionary for what all the column names mean:   [data dictionary](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-03-10/readme.md#data-dictionary)

#### Download data or Read data

```{r}
## Test if a directory named data exists locally. If it does not, write an R function that creates it programmatically. Saves the data only once
if (!file.exists(here("data", "dat_API.csv"))) {
    
    readRenviron("~/.Renviron")
    var_list <- c('B19013_001', 'B19201_001', 
                  'B08303_001', 'B08134_101', 'B08134_111',
                  'B08301_016', 'B08301_017', 'B08301_018',
                  'B08301_019', 'B08301_020', 'B09002_001')
    
    TX <- get_acs(geography = "county", 
              variables = var_list, 
              year = 2020, 
              state = 'TX',
              output = 'wide')
    
    data_filtered <- cbind(TX[,c(1,2)], TX[, !grepl('E', colnames(TX))])
    colnames(data_filtered)[3:13] = c('Mid_Income', 'Mid_nonfamily_income',
                                      'Time_toWork', 'Walk_Time_toWork',
                                      'Time_trans', 'Time_Taxicab', 'Time_Motorcycle',
                                      'Time_Bicycle', 'Time_Walked', 'Time_Other means',
                                      'Num_Child')
    
  # B19013_001: MEDIAN HOUSEHOLD INCOME
  # B19201_001: NONFAMILY HOUSEHOLD INCOME
  # B08303_001: travel time to work
  # B08134_101: Estimate!!Total:!!Walked: MEANS OF TRANSPORTATION TO WORK BY TRAVEL 
  # B08134_111: Estimate!!Total:!!Taxicab, motorcycle, bicycle, or other means:
  # B08301_016: Estimate!!Total:!!Taxicab
  # B08301_017: Estimate!!Total:!!Motorcycle
  # B08301_018: Estimate!!Total:!!Bicycle
  # B08301_019: Estimate!!Total:!!Walked
  # B08301_020: Estimate!!Total:!!Other means
  # B09002_001: Estimate!!Total: OWN CHILDREN UNDER 18 YEARS BY FAMILY

    save_directory <- here("data") # File for saving data, must be created
    if (!dir.exists(save_directory)) {
      dir.create(save_directory, recursive = TRUE)
    }
    
    # save the files to csv objects ()
    write.csv(data_filtered, file = here("data", "dat_API.csv"))
    
}
## Read in the data locally each time you knit/render
dat_API <- read.csv(here("data", "dat_API.csv")); dat_API$X = NULL
```

```{r, warning=FALSE}
library(tidyverse)
library(stringr)

# Convert the names (character strings) to all upper case.
dat_API <- dat_API %>%
  mutate(NAME = str_to_upper(NAME))

cut_point = quantile(dat_API$Mid_nonfamily_income, probs = c(0, 0.25, 0.5, 0.75, 1))

# convert Median non-family income into factor variable
dat_API$Mid_nonfamily_income <- as.factor(cut(
  dat_API$Mid_nonfamily_income,
  breaks = cut_point,
  labels = c("Low", "Medium", "High", 'Extreme High')
))

#  Start with dat_API dataset and drop any rows with NAs.
dat_API <- dat_API %>%
  drop_na()
```

```{r, warning=FALSE}
# calculate the distribution of room and board of different schooles
dat_API_summary <- dat_API %>% 
  select(-NAME) %>% 
  group_by(Mid_nonfamily_income) %>%
  summarise(
    mean_income = mean(Mid_Income),
    sd_income = sd(Mid_Income))

head(dat_API_summary)
```

```{r, warning=FALSE}
# combine the observed data and summarized result
dat_API_combined <- left_join(dat_API, dat_API_summary, 'Mid_nonfamily_income')
head(dat_API_combined)
```

```{r, warning=FALSE, fig.width=6, fig.height=4}
dat_API_combined %>% 
  ggplot(aes(x = Mid_Income, y = Time_toWork)) + 
  geom_point(aes(x = Mid_Income, y = Time_toWork),  
             linetype = "solid", color = rainbow(252), size = 1) + 
  labs(title = "Association between the Median income and Non-Family income in Texas",
       subtitle = 'Point Chart: Median income ~ Non-Family income',
       caption = "Data from US Census Bureau’s data",
       x = "Median income", y = "Non-Family income") + 
  facet_wrap(~Mid_nonfamily_income, ncol = NULL, scales = "free_y") +
  scale_x_log10()+
  scale_y_log10()+
  theme_minimal()
```

```{r, warning=FALSE, fig.width=6, fig.height=4}
dat_API_combined %>% 
  group_by(Mid_nonfamily_income) %>% 
  mutate(Mean_diff = Mid_Income - mean(Mid_Income)) %>%
  ggplot(aes(x = Mid_nonfamily_income, y = Mean_diff)) + 
  geom_histogram(stat = "identity", fill = rainbow(252)) + 
  labs(title = "The distribution of median income difference among non-family income level in Texas",
       subtitle = 'relative to the mean level at non-family income level',
       caption = "Data from US Census Bureau’s data",
       x = "State", y = "The difference compared to mean value") +
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size=8))
```

#### BKMR Model

[Using reference](https://jenfb.github.io/bkmr/overview.html)

```{r, warning=FALSE}
library(bkmr)
colnames(dat_API_combined)

data_unique = unique(dat_API_combined)
dat <- SimData(n = dim(data_unique)[1], M = 4) # introduce some nuisance data
```

#### Adopt purrr function to facilitate process

```{r}
## explore linear regression
by_degree <- split(data_unique, data_unique$Mid_nonfamily_income)
by_degree |>
  map(.f = ~ lm(Mid_Income ~ Time_toWork, data = .x)) |>
  map(.f = coef)

## transform data
data_unique[, c(2,4)] <- map_dfc(.x = data_unique[, c(2,4)], .f = tolower)
data_unique$Mid_nonfamily_income = as.integer(as.factor(data_unique$Mid_nonfamily_income))
```

```{r, warning=FALSE, fig.width=6, fig.height=4}
set.seed(208)
data_unique[,c(3:15)] = scale(data_unique[,c(3:15)])

y <- as.matrix(data_unique$Mid_Income)
Z <- cbind(as.matrix(data_unique[, c(5:12)],), dat$Z)
X <- as.matrix(data_unique[, c(13)])
fitkm <- kmbayes(y = y, Z = Z, X = X, iter = 1000, verbose = FALSE, varsel = TRUE)
ExtractPIPs(fitkm)
# A posteriori inclusion probability in a simulated data set (the bigger the better)
#TracePlot(fit = fitkm, par = "r", comp = 1)
#TracePlot(fit = fitkm, par = "sigsq.eps")

pred.resp.univar <- PredictorResponseUnivar(fit = fitkm)
library(ggplot2)
ggplot(pred.resp.univar, aes(z, est, ymin = est - 1.96*se, ymax = est + 1.96*se)) + 
    geom_smooth(stat = "identity") + 
    facet_wrap(~ variable) +
    labs(title = "The univariate relationship between specific exposure and the Median Income",
       subtitle = 'All variables have been standarized to 0~1',
       caption = "Data from US Census Bureau’s data",
       x = "Scaled Value", y = "h(z)")

```

### Summary of Analysis

-   The obvious negative linear association between Median income and Non-Family income in the Non-family income (Extreme high group), while the associations in other groups are not obvious or dominant.
-   The distribution of median income difference among non-family income level in Texas are different, and the high level group showed the lowest variation.
-   Even though we considering nuisance parameter, BKMR could identify the actual effect from Time to Work, Time of transportation and Time of Motorcycle to median income. Also, in the PIP value list, Time of transportation, Time of Motorcycle showed the top2 highest prbability.

### Functions used from packages

-   dplyr: select(); summarise(); mutate(); group_by(); left_join() <br>
-   tidyr: drop_na(); <br>
-   purrr: map_dfc(); map()
-   ggplot2: geom_point(); geom_histogram(); geom_smooth(), facet_wrap()
-   bkmr: ExtractPIPs(); kmbayes(); PredictorResponseUnivar()
